---
title: "Programming Notes"
output:
  html_document:
    highlight: textmate
    keep_md: yes
    theme: flatly
---
<hr>
## Chapter 1 
<hr>

## 1.1 The Elements of Programming

- Primitive arithmetic operations and their combinations
- Abstractions of combined operations by defining them as compound procedures

<hr>
### **1.1.7** Newton's Method for Computing Square Roots

Difference between functions and procedures:  

- A function does not describe the steps of how to go from an input to an output -- a procedure does! 
- A function describes properties of things and a procedure tells how to do things
    - Function = declarative knowledge (what is) 
    - Procedure = imperative knowledge (how to)

To compute a square root using Newton's methods,  you make iterative approximations that get better according to the calculation

    Approximation of Sqrt(y)  = y + x/y

This can be represented in an R function (or procedure)

```{r newtonfunction}

newton.sqrt <- function (x) {
    good.enough <- function(guess, x) {
        abs(guess^2 - x) < 0.001
        }
    improve.guess <- function(guess, x) {
        mean(c(guess, x/guess))
        }
    sqrt.iter <- function(guess, x) {
        if (good.enough(guess, x)) { 
            guess } else { sqrt.iter(improve.guess(guess, x), x) }
        }
    return(sqrt.iter(1, x))
    }
newton.sqrt(9)
```

Leveraging scoping rules, we can simplify the `newton.sqrt` function like so:

```{r newtonfunction.simple}

newton.sqrt.simplified <- function (x) {
    
    good.enough <- function(guess) {
        abs(guess^2 - x) < 0.001
        }
    improve.guess <- function(guess) {
        mean(c(guess, x/guess))
        }
    sqrt.iter <- function(guess) {
        if (good.enough(guess)) { 
            guess } else { sqrt.iter(improve.guess(guess)) }
        }
    return(sqrt.iter(1))
    }

newton.sqrt.simplified(9)
```

#### Lexical scoping

The functions defined within `newton.sqrt` and `newton.sqrt.simplified` are in the scope of the formal parameter `x` since it is defined in the formal arguments, so we don't have to pass `x` explicitly to each subsequent nested function. Rather, we allow a new `x` within each function to be a _free variable_ and inherit its value from the bound `x`.


If we want to compare the two functions for CPU efficiency, we can do this:

```{r computeefficiency}
system.time(newton.sqrt(123456789))
system.time(newton.sqrt.simplified(123456789))
```

Both functions illustrate another important concept: **abstraction**

### **1.1.8** Procedures as Black-box Abstractions

`sqrt.newton` is an example of a function defined by a set of mutually defined functions

Here is how `R` defines its native `sqrt` function, which calls `C++` code so that it is light and efficient:

```{r native.sqrt.function}
dummycode <- "Find the C code!"
```

Decomposition of the program into multiple inter-related procedures -- it is important that **each procedure accomplishes an identifiable task that can be used as a module in defining other procedures**

When we define the `good.enough` procedure (or "function" in `R` parlance), it takes the `square` argument (or `'^2'` function) as a 'black box'; as far as the `good.enough` procedure is concerned, `square` is not a procedure, but **an abstraction of a procedure**

The parameter names should be local to the body of the procedure

A formal parameter is a _bound variable_, and any variable within the body of the function not named in the formal arguments is a _free variable_

<hr>

## **1.2** Procedures and the Processes They Generate

<hr>

### **1.2.1** Linear Recursion and Iteration

A basic example of a function that can be computed using linear recursion and/or linear iteration is the _factorial_, defined as 

    n! = n * (n-1)!

Here it is, using linear recursion:

```{r linear.recursion}

factorial.recursive <- function(n) {
    if (n == 1) {1} else {
        n * factorial.recursive(n - 1)
        }
    }

factorial.recursive(9)
```

Factorials can also be calculated using linear iteration:

```{r linear.iteration}
factorial.iterative <- function(n) {
    fact.iter <- function(product, counter) {
        if (counter > n) { product } else {
          fact.iter(
              (counter * product),
              (counter + 1)
              )}
        }
    fact.iter(1, 1)
    }
factorial.iterative(9)
```

(As an aside, here is a comparison of how they perform)

```{r compfunctions}
system.time(factorial.recursive(100))
system.time(factorial.iterative(100))
```


